Water, High-Altitude Condensates, and Possible Methane Depletion in the Atmosphere of the Warm Super-Neptune WASP-107\lowercase{b}

### 2018-04-17 ApJL_submission/ms.tex
final response 3

\author{Laura Kreidberg}
\affiliation{Harvard Society of Fellows\
78 Mt. Auburn St.\\
Cambridge, MA 02138, USA}
\affiliation{Harvard-Smithsonian Center for Astrophysics\
60 Garden St.\\
Cambridge, MA 02138}
\nocollaboration

\author{Michael R. Line}
\affiliation{Arizona State University}
\nocollaboration

\author{Daniel Thorngren}
\affiliation{University of California Santa Cruz}
\nocollaboration

\author{Caroline V. Morley}
\altaffiliation{Sagan Fellow}
%\affiliation{Harvard-Smithsonian Center for Astrophysics}
\affiliation{Harvard-Smithsonian Center for Astrophysics\
60 Garden St.\\
Cambridge, MA 02138}
\nocollaboration

\author{Kevin B. Stevenson}
\affiliation{Space Telescope Science Institute}
\nocollaboration

\begin{abstract}
	The super-Neptune exoplanet WASP-107b is an exciting target for atmosphere characterization. It has an unusually large atmospheric scale height and a small, bright host star, raising the possibility of precise constraints on its current nature and formation history.  We report the first atmospheric study of WASP-107b, a Hubble Space Telescope measurement of its near-infrared transmission spectrum.  We determined the planet's composition with two techniques: atmospheric retrieval based on the transmission spectrum and interior structure modeling based on the observed mass and radius. The interior structure models set a $3\,\sigma$ upper limit on the atmospheric metallicity of $30\times$ solar. The transmission spectrum shows strong evidence for water absorption ($6.5\,\sigma$ confidence), and the retrieved water abundance is consistent with expectations for a solar abundance pattern. \added{The inferred carbon-to-oxygen ratio is subsolar at $2.7\,\sigma$ confidence, which we attribute to possible methane depletion in the atmosphere.}  The spectral features are smaller than predicted for a cloud-free composition, crossing less than one scale height. A thick condensate layer at high altitudes (0.1 - 3 mbar) is needed to match the observations. \added{We find that physically motivated cloud models with moderate sedimentation efficiency ($f_\mathrm{sed} = 0.3$) or hazes with a particle size of 0.3 $\mu$m reproduce the observed spectral feature amplitude.} Taken together, these findings serve as an illustration of the diversity and complexity of exoplanet atmospheres. The community can look forward to more such results with the high precision and wide spectral coverage afforded by future observing facilities. 
\end{abstract}

@@

\begin{figure*}
\includegraphics[width = \textwidth]{fig2_spectrum.pdf}
\caption{The transmission spectrum of WASP-107b (points with 1\,$\sigma$ error bars) compared to retrieved models (blue line with shaded $1$ and $2\sigma$ confidence intervals). The features at $1.15$ and $1.4$ $\mu$m are due to water absorption. The right-hand axis indicates the normalized atmospheric scale height assuming solar composition. The water feature crosses less than one scale height, indicating that condensates are present at high altitude.} 
\label{fig:spectrum}
\end{figure*}

@@

We explored a few retrieval scenarios with additional complexity, including the addition of cloud patchiness \citep{line16}, a quench pressure for nitrogen and carbon species \citep[e.g.][]{morley17}, and a power law haze opacity.  We also varied the assumed planet mass by $3\,\sigma$. These models did not significantly improve the fit quality, and the main retrieval results were unchanged.  We also explored cases without the $30\times$ solar metallicity prior upper limit. Without this prior, \added{the retrieved metallicity is $[M/H] = 1.35^{+0.68}_{-1.48}$, log10(C/O) is $-1.57^{+0.54}_{-0.31}$, and cloud top pressure $\log P_\mathrm{c} = -3.63^{+0.81}_{-0.48}.$ These results show that the inferred C/O ratio and cloud-top pressure are unaffected by the prior on atmospheric metallicity.  The metallicity is consistent with the interior structure modeling (which puts a $3\,\sigma$ upper limit on metallicity of $[\mathrm{M/H}] = 1.5$). The transmission spectrum does not constrain the atmospheric metallicity more precisely because there is a degeneracy between metallicity and the planet's 10-bar radius \citep[e.g]{griffith14,heng17} with the constraint primarily driven by the H$_2$O-H$_2$ CIA continuum ratio, molecular weight \citep{line16}, and pressure broadening \citep{dewit13}.}

>>>>

\author{Laura Kreidberg}
\affiliation{Harvard Society of Fellows, 78 Mt. Auburn St., Cambridge, MA 02138, USA}
\affiliation{Harvard-Smithsonian Center for Astrophysics, 60 Garden St., Cambridge, MA 02138}

\author{Michael R. Line}
\affiliation{School of Earth \& Space Exploration, Arizona State University, Tempe AZ 85287, USA}

\author{Daniel Thorngren}
\affiliation{Department of Astronomy and Astrophysics, University of California, Santa Cruz, CA 95064}

\author{Caroline V. Morley}
\altaffiliation{Sagan Fellow}
\affiliation{Harvard-Smithsonian Center for Astrophysics, 60 Garden St., Cambridge, MA 02138}

\author{Kevin B. Stevenson}
\affiliation{Space Telescope Science Institute, 3700 San Martin Drive, Baltimore, MD 21218, USA}

\begin{abstract}
	The super-Neptune exoplanet WASP-107b is an exciting target for atmosphere characterization. It has an unusually large atmospheric scale height and a small, bright host star, raising the possibility of precise constraints on its current nature and formation history.  We report the first atmospheric study of WASP-107b, a Hubble Space Telescope measurement of its near-infrared transmission spectrum.  We determined the planet's composition with two techniques: atmospheric retrieval based on the transmission spectrum and interior structure modeling based on the observed mass and radius. The interior structure models set a $3\,\sigma$ upper limit on the atmospheric metallicity of $30\times$ solar. The transmission spectrum shows strong evidence for water absorption ($6.5\,\sigma$ confidence), and the retrieved water abundance is consistent with expectations for a solar abundance pattern. The inferred carbon-to-oxygen ratio is subsolar at $2.7\,\sigma$ confidence, which we attribute to possible methane depletion in the atmosphere.  The spectral features are smaller than predicted for a cloud-free composition, crossing less than one scale height. A thick condensate layer at high altitudes (0.1 - 3 mbar) is needed to match the observations. We find that physically motivated cloud models with moderate sedimentation efficiency ($f_\mathrm{sed} = 0.3$) or hazes with a particle size of 0.3 $\mu$m reproduce the observed spectral feature amplitude. Taken together, these findings serve as an illustration of the diversity and complexity of exoplanet atmospheres. The community can look forward to more such results with the high precision and wide spectral coverage afforded by future observing facilities. 
\end{abstract}

@@

\begin{figure*}
\includegraphics[width = \textwidth]{fig2_spectrum.pdf}
\caption{The transmission spectrum of WASP-107b (points with 1\,$\sigma$ error bars) compared to retrieved models (blue line with shaded $1$ and $2\sigma$ confidence intervals). The features at $1.15$ and $1.4$ $\mu$m are due to water absorption. The right-hand axis indicates the normalized atmospheric scale height assuming solar composition. The water feature crosses less than one scale height, indicating that condensates are present at high altitude. \added{We also show the best fit model from an analysis where the C/O is fixed at 0.54, the solar value (dashed line). There are subtle differences in feature shape due to methane absorption in the solar C/O case, which cause the retrieval to favor sub-solar C/O values at 2.7$\sigma$ confidence.}} 
\label{fig:spectrum}
\end{figure*}

@@

We explored a few retrieval scenarios with additional complexity, including the addition of cloud patchiness \citep{line16}, a quench pressure for nitrogen and carbon species \citep[e.g.][]{morley17}, a power law haze opacity, \added{no prior on atmospheric metallicity, a more flexible T/P profile (with added free parameters for infrared opacity and visible-to-infrared opacity), and the more flexible T/P profile with quenching}.  We also varied the assumed planet mass by $3\,\sigma$. These models did not significantly improve the fit quality, and the \added{retrieved cloud-top pressure and chemical composition were consistent to within $0.1\,\sigma$}.

\added{For example, when we removed the $30\times$ solar metallicity prior upper limit,} the retrieved metallicity is $[M/H] = 1.35^{+0.68}_{-1.48}$, log10(C/O) is $-1.57^{+0.54}_{-0.31}$, and cloud top pressure $\log P_\mathrm{c} = -3.63^{+0.81}_{-0.48}.$ These results show that the inferred C/O ratio and cloud-top pressure are unaffected by the prior on atmospheric metallicity.  The metallicity is consistent with the interior structure modeling (which puts a $3\,\sigma$ upper limit on metallicity of $[\mathrm{M/H}] = 1.5$). The transmission spectrum does not constrain the atmospheric metallicity more precisely because there is a degeneracy between metallicity and the planet's 10-bar radius \citep[e.g]{griffith14,heng17} with the constraint primarily driven by the H$_2$O-H$_2$ CIA continuum ratio, molecular weight \citep{line16}, and pressure broadening \citep{dewit13}.

\added{Similarly, we find that the transmission spectrum does not provide a strong constraint on the shape of the temperature pressure profile. The primary constraint on temperature structure is the effective scale height temperature, which sets the amplitude of features in the spectrum. To confirm this, we explored using a more flexible T/P profile with free parameters for infrared and visible-to-infrared opacity, in addition to irradiation temperature (which we label case 2). Compared to our nominal model (case 1), which just varies irradiation temperature but keeps the T/P profile shape fixed, we found that the retrieved temperature at the photosphere is nearly identical. For example, at $P = 10^{-4}$ bar (the most probable cloud-top pressure), we find $T = 618^{+150}_{-120}$ K for case 1, and $T = 615^{+178}_{-123}$ K for case 2. The constraints on chemical composition agree at the 0.1-$\sigma$ level for the two cases.}
